### 2017-11-06 paper/sections/1_introduction.tex
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and facilitating rapid
  development. Changing the policy in use is critical in applications such as
  ParSplice that have alternating stable and chaotic simulation ``access regimes"
  over the course of a long-running simulation.

  \item has useful primitives that, while designed for file systems, turn
  out to also be effective for cache management. This finding shows how the Mantle
  engine generalizes to a different domain and code-base.

@@

@@

% this gives us many policies that are effective across disciplines
% - reuse: eases burden of writing policies
% - autonomic: lays groundwork for an adaptable policy that mixes/matches policies

This last contribution is explored in Section~\S\ref{sec:mantle-brains}, where
we try a range of policies from different disciplines; but more importantly, in
Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
of policies we designed for ParSplice's cache management are very similar to
the policies in the Ceph file system that are used to load balance metadata.
This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Such work is made possible with Mantle's ability to dynamically change
policies.

>>>>

\begin{abstract}

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the Mantle
policy engine to dynamically explore a variety of techniques, ranging from
basic algorithms and heuristics to statistical models, calculus, and machine
learning. While Mantle was originally designed for distributed file systems, we
show how it effectively decomposes the problem into manageable policies for a
different domain and service (in this case, cache management).  Our exploration
of this space results in a two policy scheme that achieves 96\% efficiency
while using only 7.6\% of the memory resources required by the base case.

\end{abstract}

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and facilitating rapid
  development. Changing the policy in use is critical in applications such as
  ParSplice that have alternating stable and chaotic keyspace access patterns
  over the course of a long-running simulation.

@@

  \item has useful primitives that, while designed for file system metadata
  load balancing, turn out to also be effective for cache management. This
  finding shows how the policy engine generalizes to different domains and
  enables control of policies by machines instead of administrators.

@@

% this gives us many policies that are effective across disciplines
% - reuse: eases burden of writing policies
% - autonomic: lays groundwork for an adaptable policy that mixes/matches policies
% FUTURE WORK

This last contribution is explored in Section~\S\ref{sec:mantle-brains}, where
we try a range of policies from different disciplines; but more importantly, in
Section~\S\ref{sec:relation-to-file-systems}, we conclude that the collection
of policies we designed for ParSplice's cache management are very similar to
the policies in the Ceph file system that are used to load balance metadata,
suggesting that there is potential for automatically adapting and generating
policies dynamically.

### 2017-11-06 paper/sections/2_parsplice-keyspace-analysis.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{ParSplice Keyspace Analysis}

\subsection{Structured Access Regimes}

Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%shows ``access regimes", or groups of accesses
%to the same subset of keys:
\begin{itemize}
  \item change immediately, different sizes
  \item monotonically increasing
  \item random access to a single key
  \item early keys accessed more
\end{itemize}

\noindent\textbf{Conclusion}: unique patterns of a real HPC application

\subsection{Cache Size Analysis}

Figure~\ref{fig:keyspace-analysis_cachesize}.

\noindent\textbf{Conclusion}: workload phases (high request rate to a small
number of keys and then low request rate to a large number keys) need a dynamic
load balancing policy.

\subsection{Overfitting Policies}

Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.

\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
sensitive to: max usage or usage over time.\\

\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
request rate for a 2.5 hour run, but it is infeasible to do this for every
combination of system setup, job lengths, parsplice parameters.

\subsubsection*{Discussion}
Why don't we just an LRU cache:
\begin{itemize}
  \item we can do better (workload is structured and has locality)
  \item finding the size of the cache is hard
  \item hotspots dissipate too quickly
\end{itemize}

>>>>

%\section{ParSplice Keyspace Analysis}
%
%We demonstrate these contributions by changing the input to our molecular
%dynamics simulation, which changes the keyspace access patterns. 
%
%\subsection{Structured Access Regimes}
%
%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%%shows ``access regimes", or groups of accesses
%%to the same subset of keys:
%\begin{itemize}
%  \item change immediately, different sizes
%  \item monotonically increasing
%  \item random access to a single key
%  \item early keys accessed more
%\end{itemize}
%
%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
%
%\subsection{Cache Size Analysis}
%
%Figure~\ref{fig:keyspace-analysis_cachesize}.
%
%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
%number of keys and then low request rate to a large number keys) need a dynamic
%load balancing policy.
%
%\subsection{Overfitting Policies}
%
%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
%
%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
%sensitive to: max usage or usage over time.\\
%
%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
%request rate for a 2.5 hour run, but it is infeasible to do this for every
%combination of system setup, job lengths, parsplice parameters.
%
%\subsubsection*{Discussion}
%Why don't we just an LRU cache:
%\begin{itemize}
%  \item we can do better (workload is structured and has locality)
%  \item finding the size of the cache is hard
%  \item hotspots dissipate too quickly
%\end{itemize}

\section{ParSplice Keyspace Analysis}
\label{sec:parsplice-keyspace-analysis}

ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
(MD) simulation package developed at LANL. It is part of the Exascale Computing
Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
Materials for the Future initiative. Its phases are:

\begin{enumerate}

  \item a splicer tells workers to generate segments (short MD trajectory) for
  specific states

  \item workers read initial coordinates for their assigned segment from data
  store; the key-value pair is (state ID, coordinate)

  \item upon completion, workers insert final coordinates for each segment into
  data store, and wait for new segment assignment

\end{enumerate}

The computation can be parallelized by adding more workers or by adding worker
tasks to parallelize individual workers.  The workers are stateless and read
initial coordinates from the data store each time they begin generating
segments. Since worker tasks do not maintain their own history, they can end up
reading the same coordinates repeatedly. To mitigate the consequences of these
repeated reads, ParSplice provisions a hierarchy of nodes to act as caches that
sit in front of a single node persistent database.  Values are written to each
tier and reads traverse up the hierarchy until they find the data.

We use ParSplice to simulate the evolution of metallic nanoparticles that grow
from the vapor phase. As the run progresses, the energy landscape of the system
becomes more complex. Two domain factors control the number of entries in the
data store: the growth rate and the temperature. The growth rate controls how
quickly new atoms are added to the nanoparticle: fast growth rates lead to
non-equilibrium conditions, and hence increase the number of states that can be
visited. However, as the particle grows, the simulation slows down because the
calculations become more expensive, limiting the rate at which new states are
visited. On the other hand, the temperature controls how easily a trajectory
can jump from state to state; higher temperatures lead to more frequent
transitions.

The nanoparticle simulation stresses the data store architecture of ParSplice.
It visits more states than other input decks because the system uses a cheap
potential, has a small number of atoms, and operates in a complex energy
landscape with many accessible states. Changing growth rates and temperature
alters the size, shape, and locality of the data store keyspace. Lower
temperatures and smaller growth rates create hotter keys with smaller keyspaces
as many segments are generated in the same set of states before the trajectory
can escape to a new region of state space.

Our evaluation uses the total ``trajectory length" as the goodness metric. This
value is the duration of the overall trajectory produced by ParSplice. At
ideal efficiency, the trajectory length should increase with the square root of
the wall-clock time, since the wall-clock cost of time-stepping the system by
one simulation time unit increases in proportion of the total number of atoms.

\subsection{ParSplice Keyspace Analysis}
\label{sec:parsplice-keyspace-analysis}
%IT HAS 8K keys!  How did we take these measurements

We instrumented ParSplice with performance counters and keyspace counters.  The
performance counters track ParSplice progress while keyspace counters track
which keys are being accessed by the ParSplice ranks. Because the keyspace
counters have high overhead we only turn them on for the keyspace analysis.
The cache hierarchy is unmodified but for the persistent database node, we
replace BerkeleyDB on NFS with LevelDB on Lustre. Original ParSplice
experiments showed that BerkeleyDB's syncs caused reads/writes to bottleneck on
the persistent database. We also use Riak's customized
LevelDB\footnote{https://github.com/basho/leveldb} version, which comes
instrumented with its own set of performance counters.

All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
cores per node.  Each node has 128GB of RAM and our goal is to limit the size
of the cache to 3\% of RAM\footnote{Empirically, this is a threshold that we
find to work well for most applications}. Note that this is an addition to the
30GB that ParSplice uses to manage other ranks on the same node.  A single Cray
node produced trajectories that are \(5\times\) times longer than our 10 node
CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As
a result, it reaches different job phases faster and gives us a more
comprehensive view of the workload. The performance gains compared to the
commodity clusters have more to do with memory/PCI bandwidth than network.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keyspace.png}\\
  \caption{The keyspace size is small but must satisfy many reads as workers
  calculate new segments. It is likely that we will need
  more than one node to manage segment coordinates when we scale the system or jobs up.
  \label{fig:methodology-keyspace}}
\end{figure}

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-keys.png}\\
  \caption{The keyspace imbalance is due to workers generating deep
  trajectories and reading the same coordinates. Over time, the accesses get
  dispersed across different coordinates resulting in some keys being more
  popular than others.\label{fig:methodology-keys}}
\end{figure}

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (black annotations) and request load (bars) after a one hour run
with a different number of workers (\(x\) axis). While the keyspace size and
capacity is relatively modest the memory usage scales linearly with the number
of workers. This is a problem if we want to scale to Trinitite's 6000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 20GB for the cache.  This memory utilization easily eclipses
the 3\% threshold we set earlier, even without factoring in the memory usage
from other workers.

\subsubsection*{An active but small keyspace} The bars in
Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
worker tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when worker tasks generate segments for the same
state, which is computationally cheap (this motivates
Section~\S\ref{sec:static-load-balancing}).

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:the-need-for-dynamic-load-balancing-policies}).  The growth
rate, temperature, and number of workers have a predictable effect on the
structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different growth rates, but that the spatial
locality of key accesses is similar ({\it e.g.}, some keys are still
\(5\times\) more popular than others).  Figure~\ref{fig:motivation-regimes}
shows how entropy for different growth rates has temporal locality, as the
reads per second for \(\Delta_2\) looks like the reads per second for
\(\Delta_1\) stretched out along the time axis.  Trends also exist for
temperature and number of workers but are omitted here for space. This
structure means that we can learn the regimes and adapt the storage system
(this motivates Section~\S\ref{sec:ml-for-the-keyspace}).

### 2017-11-06 paper/sections/6_conclusion.tex
paper: switch to IEEE template, hack on intro, move future work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

>>>>

\section{Future Work}

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Such work is made possible with Mantle's ability to dynamically change
policies.
